\thispagestyle{empty}
\chapter{Acknowledgements}
\large{

This Master thesis has been written within the Erasmus Mundus Master's program \textit{Computer Simulations for Science and Engineering (COSSE)}\footnote{For more information, please visit: \url{http://www.kth.se/en/studies/programmes/em/cosse}}. During the past two years of my studies, I had the opportunity to study for one year at the Royal Institute of Technology in Stockholm, Sweden, as well as the Technical University in Delft, the Netherlands. But COSSE offers much more than the possibility to study at two world-known universities in the field of Computer Science and Applied Mathematics and, therefore, I am mostly thankful for the friendship and scientific influence of all the great people I met from all over the world.

Since the beginning of my studies at TU Berlin more than five years ago, I was also supported by a scholarship of the \textit{Friedrich-Ebert foundation (FES)}\footnote{Again, more information is available online: \url{http://www.fes.de/studienfoerderung}}. The FES is a political foundation that mostly supports students from a so-called working class background. During my time as a stipend, I went to many interesting workshops and seminars, especially in the field of development politics.

I would like to thank professor Volker Mehrmann and professor G\"unter B\"arwolff from TU Berlin who supported my application for COSSE. Because of them I felt encouraged to apply for an international study program and the Erasmus Mundus scholarship. Professor Mehrmann also gave crucial scientific advices to my thesis work whenever I was in Berlin.

Most of all, I would like to thank the (daily) supervisors of my Master thesis, Martin van Gijzen and Marielba Rojas. Our weekly meetings, known as the \textit{3M meetings}, always were inspiring, encouraging and, on a personal level, delightful. Both of you motivated me to continue my scientific career with a PhD. Moreover, I would like to thank S\l awomir Szklarz who helped me in many respects since I arrived in Delft.

During the last two years of studying abroad, the support of my parents never ended and I am very thankful for many of their advices during our \textsc{Skype} talks. Last but not least, I would like to thank \textit{Jana}. Only because of you, I never felt lonely away from home.
} 